---
title: "p8105_hw1_zn2220"
author: "Ziang Niu"
date: "2025-09-19"
output: github_document
---

```{r setup}
library(moderndive)
library(ggplot2)
library(tidyverse)
```

# Problem 1

```{r echo=FALSE}
data("early_january_weather")
```

Here is a summary of the dataset, including the names and some indicators of the dataset.

```{r}
summary(early_january_weather)
```

It has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns.

Some head data can be shown here.

```{r echo=FALSE}
head(early_january_weather)
```

The mean temperature is `r mean(early_january_weather$temp, na.rm = TRUE)` °F.

And the plot:

```{r echo=FALSE}
sp <- ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point(alpha = 0.75) +
  labs(
    title = "Temperature vs Time",
    x = "Time (hourly)",
    y = "Temperature (°F)",
    color = "Humidity"
  ) +
  theme_minimal()

sp
```

The temperature rose overall, from a range of 25-40°F at the beginning of January to above 50°F around January 14th, before decreasing. There were also clear daily fluctuations, with a maximum and minimum temperature difference of approximately 5-15°F, especially during the first week. Humidity also increased, as indicated by the plot's points: the dark points on the left represent low humidity, while the lighter points on the right represent high humidity. Temperature and humidity are positively correlated, as the humidity rises with the temperature. On January 14th, both the temperature and humidity reached their peaks simultaneously.

Save it!
```{r echo=FALSE}
ggsave("early_jan_weather_scatter.png", plot = sp)
```

# Problem 2

Create the data frame!

```{r echo=FALSE}
set.seed(3)

normal_sample <- rnorm(10)

logical_vector <- normal_sample > 0

char_vector <- letters[1:10]

factor_vector <- factor(sample(c("Level1", "Level2", "Level3"), 10, replace = TRUE))

df <- data.frame(
  NormalSample = normal_sample,
  LogicalVector = logical_vector,
  CharVector = char_vector,
  FactorVector = factor_vector
)

df
```

Trying to calculate the means of these data.

```{r echo=FALSE, results='hide'}
NS <- mean(pull(tibble(df), NormalSample))
LV <- mean(pull(tibble(df), LogicalVector))
CV <- mean(pull(tibble(df), CharVector))
FV <- mean(pull(tibble(df), FactorVector))
```

It shows that the mean of NormalSample is `r NS` and LogicalVector is `r LV`. According to warnings, it fails to calculate the mean of CharVector and FactorVector.

Let's turn them to numeric!

```{r results='hide'}
char_num <- as.numeric(factor(char_vector))

char_num

factor_num <- as.numeric(factor_vector)

factor_num
```

Boolean variables are automatically converted to numeric in R, so their mean can be calculated directly. However, character and factor variables cannot. For characters, we need to define a mapping (for example, by converting them to factors), and factors already have internal codes, so we can apply the as.numeric() function directly.